Word-Referent Identification Under Multimodal Uncertainty

### 2018-04-15 Writeup/ms.Rmd
modif to ms

How does the listener deal with uncertainty to identify the speaker's intended word? One rigorous way to approach this question is through conducting an *ideal observer* analysis. This research strategy provides a characterization of the task/goal and shows what the optimal performance should be under this characterization.^[It is, thus, a general instance of the rational approach to cognition [@anderson90]. It can also be seens as an instance Marr's computational level of analysis.] When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. When the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. Thus, it is supposed to be used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003; @rahnev2018]. When there is a deviation from the ideal, it can reveal extra constraints on human cognition, such as limitations on the working memory or attentional resources.

@@

On the face of it, the question of combining information from the sound and the visual referent might seem similar to that of audio-visual speech integration. Nevertheless, there are at least two fundamental differences between these two cases, and both can influence the way the auditory and visual cues are integrated:

\noindent First, in the case of audio-visual speech, both modalities offer information about the same underlying speech category. They may differ only in terms of their informational reliability. In a referential context, however, the auditory and visual modalities are additionally different in terms of the roles they play in the referential process: the auditory input represents the *symbol* whereas the visual input represents the *meaning*. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].^[There is a debate as to whether speech is privileged for infant and adults for the same reasons. Whereas some researchers suggest that speech is privileged for both infants and adults because of its ability to refer [e.g., @waxman1995], others suggest that speech might *not* have a referential status from the start. Rather, speech might be prefered by infants only because of a low level auditory ``overshadowing'' [e.g., @sloutsky2003].] Thus, in a referential context, it is possible that listeners do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a sub-optimal bias for the auditory modality beyond what is expected from informational reliability alone.

@@

We tested the ideal observer model's predictions in three behavioral experiments where we varied the source of uncertainty. In Experiment 1, audio-visual tokens were ambiguous with respect to their category membership only. In Experiment 2, we intervened by adding background noise to the auditory modality, and in Experiment 3, we intervened by adding background noise to the visual modality. In all experiments, participants were quantitatively near-optimal, though overall response precision was slightly lower than expected. Moreover, in Experiment 1 where neither of the modalities was perturbed with background noise, participants weighed auditory and visual cues according to the relative reliability predicted by the optimal model. In other words, we found no evidence for a modality bias towards either the auditory or the visual modality. However, in Experiment 2 and 3, participants over-relied on one modality when the other modality was perturbed with additional noise.

@@

We construct an ideal observer model that combines probabilistic information from the auditory and visual modalities. In contrast to the  model used in most research on multisensory integration [e.g., @ernst02]---which typically studies continuous stimuli (e.g., size, location)---this probabilistic information cannot be characterized with *sensory noise/variability*, only.  Indeed, our task involves responses over categorical variables (phonemes and concepts), and therefore, the optimal model in our case should take into account, not only the noise variability around an individual perceptual estimate, but also its *within category variability*, i.e., the uncertainty related to whether this perceptual estimate belongs to a given category [see also @Bankieris17; @bejjanki2011]. In what follows, we describe a model that accounts for both type of variability. First, we describe the model in the simplified case of categorical variability only. Second, we augment this simplified model to account for sensory noise.

@@

The predictions of the optimal model will be compared to two baselines. The first baseline is a visual model which assumes that participants rely only on visual information, and an auditory model, which assumes that participants rely only on auditory information. More precisely, these baseline models assumes that the participants' responses in the bimodal condition will not be different from their response in either the visual-only or the auditory-only condition. If, as we expect, the participants rely on both the auditory and the visual modalities to make decision in the bimodal condition, the optimal model would explain more variance in human responses than the visual or the auditory model do.

@@

The optimal model (as well as the auditory and visual baselines) are *normative* models. Their predictions are made about human data in the bimodal condition, but their crucial parameters (i.e., variances associated with the visual and auditory modalities) are derived from data in the unimodal conditions.
In addition to these normative models, we consider a *descriptive* model. The parameters of this model are fit to actual responses in the bimodal condition. If the referential task induces sub-optimality (due, for instance, to the arbitrary nature of the sound-object association), then the descriptive model should explain more variance than the optimal model does.

@@

There are three possible ways human responses can deviate from optimality. These scenarios are illustrated in Figure\ \@ref(fig:subOptim) and are as follows:

@@

The percentage of participants who relied exclusively on either the visual modality or the auditory modality was `r 100*(only_visual_N_3+only_sound_N_3)/(only_visual_N_3+only_sound_N_3+both_N_3)`%, which is closer to the percentage of Experiment 2, except that now almost all of them relied on the auditory modality (`r 100*(1-exlusive_prop_3)`%). For ease of comparison, Table \@ref(tab:exclusive) provides a summary of the numbers across the three Experiments.

@@

When identifying a spoken word under uncertainty, one often needs to make the most of the available cues. Some previous work studied optimal behavior under uncertainty from the auditory input only [e.g., @clayard08; @feldman2009], and others studied optimality under multimodal uncertainty in auditory speech and visual facial features [e.g. @bejjanki2011]. The current work explored, for the first time, the case of word identification under uncertainty in speech (word form) and the visual word *referent*. More specifically, we conducted an ideal observer analysis of the task whereby a model provided predictions about how information from each modality should be combined in an optimal fashion. The predictions of the model were tested in a series of three experiments where instances of both the form and the meaning were ambiguous with respect to their category membership only (Experiment 1), when instances of the form were perturbed with additional background noise (Experiment 2), and when instances of the referent were perturbed with additional visual noise (Experiment 3).

@@

The second case of sub-optimality is related to how participants weighed the cues from the visual and the auditory modalities in a noisy context. In contrast to Experiment 1 where the combination was indistinguishable from the optimal prediction, results of Experiment 2 and 3 which both involved background noise in one modality, showed that participants had a systematic preference for the other (non-noisy) modality. From previous empirical studies, we know that when the speech signal is degraded, people tend to compensate by relying more on other sources of information such as the accompanying visual cues (i.e., lip movements) or the semantic/syntactic context [see @mattys12 for a review]. However, and generally speaking, these studies do not differentiate between an optimal compensatory strategy (i.e., relying more on the alternative source while using all information still available in the distorted signal), and a sub-optimal strategy (i.e., relying more on the alternative source while ignoring at least some of the information still available in the distorted signal). The formal approach followed in this paper allowed us to tease apart these two possibilities, and the analysis supports the sub-optimal compensatory strategy: The preference for the non-noisy modality is above and beyond what can be explained by the relative reliability alone, meaning that the participants tend to ignore at least part of the information still available in the noisy modality.

@@

Though we only tested adults in this paper, the problem of word recognition under uncertainty, as well as the need to make the most of ambiguous cues, is a particularly pressing issue for children. In fact, whereas adults are mostly faced with uncertainty in the *input*, children have to deal with the additional uncertainty that results from their early unrefined *representations* of both phonological and semantic categories. For example, upon hearing a noisy instance of "bee", adults may have to decide whether the speaker intended to say "pea" or "bee", but children can additionally be uncertain whether "bee" is a different word from "pee" (as opposed to, say, a valid within category variation), especially if these similar sounding words are newly learned [@stager1997; @Merriman91; @Creel2012; @Swingley2016; @white2008b]. Though new word form representation can be shown to be differentiated under some circumstances [e.g., @yoshida2009], this differentiation is still not mature enough and is probably noisier than the adult-like representation and/or encoded with lower confidence [see @Swingley2007].

@@

Thus, uncertainty in the representation associated with one modality (e.g., a bee and a fly) can be mitigated through the possibly more differentiated representations associated with the other modality (e.g., the sound "bee" is acoustically different from the sound "fly"). That being said, a multi-modal cue combination strategy might help children not only recognize an individual word instance, but also refine the underlying phonological and semantic representations in the process. Previous research in early word learning has---whether implicitly or explicitly---largely treated the process of learning form and of learning meaning as independent. However, the developmental data reviewed above shows that children do not wait to have completed the acquisition of form to start learning meanings, and that both form and meaning representations develop, rather, in a parallel fashion. A few studies pointed to the possibility of an interaction between sound and meaning in early acquisition. For instance, @waxman1995 showed that labeling various objects with the same name helps infants form the broad semantic category [but see @sloutsky2003]. Vice versa, @yeung09 showed that pairing similar sounds with different objects help infants pay attention to subtle phonological contrasts. The present study proposes a first step towards a formal framework where accounts of sound-meaning synergy in development can be unified and further explored. For example, one could imagine that, initially, visual and auditory categories have relatively large noise variances, and that development consists in reducing the values of the variances through a mutually constraining process as further multimodal data accumulates.

One limitation of the work is that we used simplified stimuli. For the auditory modality, we used speech categories that varied along a single acoustic dimension. While this dimension might be sufficient to recognize  words in our specific case, in general the speech signal may be more complex, varying along several acoustic/phonetic dimensions. Additionally, these dimensions may be highly variable due to various kinds of speaker and context differences. The same thing can be said about the referential stimuli. Here we used a continuum along a single morph dimension in order to construct a multimodal input where the auditory and visual components have symmetrical properties. Though such morph is not the exact visual variability that people would encounter in they daily lives, it allowed us to precisely test the role of auditory and visual information in the cue combination process. Parameterizing semantic dimensions is a notoriously difficult problem, but morphs have been used in previous research as a reasonable proxy [@freedman2001; @havy2016; @sloutsky2004]. 
It is an open question as to whether people use the same strategy in controlled laboratory conditions, as in more naturalistic settings where they have to deal with various levels of variability. However an answer to this question is likely to involve a multifaceted approach, involving---besides laboratory experiments---analyses of corpora with a more realistic multimodal input [e.g., @fourtassi2014b; @harwath2016; @roy2015].

In conclusion, this work studied the mechanism of word identification under uncertainty in both the word form and the word referent. To our knowledge, this is the first study that performs an ideal observer analysis of this task. We found people to be near optimal in their cue combination: They weighted each modality according to its relative reliability. However, they also showed patterns of sub-optimality especially when the stimuli were perturbed with additional background noise. Though the present study did not directly address the issue of early word learning, it provides a framework where developmental questions can also be investigated. For instance, future work should explore whether children, like adults, use probabilistic cues from both the auditory and the visual input to recognize ambiguous words,  the extent to which they combine these cues in an optimal fashion, and whether these combination help them with refining their early phonological and semantic representations.

>>>>

How does the listener deal with uncertainty to identify the speaker's intended word? One rigorous way to approach this question is through conducting an *ideal observer* analysis. This research strategy provides a characterization of the task/goal and shows what the optimal performance should be under this characterization.^[It is, thus, a general instance of the rational approach to cognition [@anderson90]. It can also be seens as an instance of Marr's computational level of analysis.] When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. When the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. It is supposed to be used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003; @rahnev2018]. When there is a deviation from the ideal, it can reveal extra constraints on human cognition, such as limitations on the working memory or attentional resources.

@@

On the face of it, the question of combining information from the sound and the visual referent might seem similar to that of audio-visual speech integration. Nevertheless, there are at least two fundamental differences between these two cases, and both can influence the way the auditory and visual cues are combined:

\noindent First, in the case of audio-visual speech, both modalities offer information about the same underlying speech category. They may differ only in terms of their informational reliability. In a referential context, however, the auditory and visual modalities are additionally different in terms of the roles they play in the referential process: the auditory input represents the *symbol* whereas the visual input represents the *meaning*. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].^[There is a debate as to whether speech is privileged for children and adults for the same reasons. Whereas some researchers suggest that speech is privileged for both children and adults because of its ability to refer [e.g., @waxman1995], others suggest that speech might *not* have a referential status from the start. Rather, speech might be prefered by children only because of a low level auditory ``overshadowing'' [e.g., @sloutsky2003].] Thus, in a referential context, it is possible that listeners do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a sub-optimal bias for the auditory modality beyond what is expected from informational reliability alone.

@@

We tested the ideal observer model's predictions in three behavioral experiments where we varied the source of uncertainty. In Experiment 1, audio-visual tokens were ambiguous with respect to their category membership only. In Experiment 2, we intervened by adding background noise to the auditory modality, and in Experiment 3, we intervened by adding background noise to the visual modality. In all experiments, participants were quantitatively near-optimal, though overall response precision was slightly lower than expected. Moreover, in Experiment 1 where neither of the modalities was perturbed with background noise, participants weighted auditory and visual cues according to the relative reliability predicted by the optimal model. In other words, we found no evidence for a modality bias towards either the auditory or the visual modality. However, in Experiment 2 and 3, participants over-relied on one modality when the other modality was perturbed with additional noise.

@@

We construct an ideal observer model that combines probabilistic information from the auditory and visual modalities. In contrast to the  model used in most research on multisensory integration [e.g., @ernst02]---which typically studies continuous stimuli (e.g., size, location)---this probabilistic information in our case cannot be characterized with *sensory noise/variability*, only.  Indeed, our task involves responses over categorical variables (phonemes and concepts), and therefore, the optimal model in our case should take into account, not only the noise variability around an individual perceptual estimate, but also its *categorical variability*, i.e., the uncertainty related to whether this perceptual estimate belongs to a given category [see also @Bankieris17; @bejjanki2011]. In what follows, we describe a model that accounts for both type of variability. First, we describe the model in the simplified case of categorical variability only. Second, we augment this simplified model to account for sensory noise.

@@

The predictions of the optimal model will be compared to two baselines. The first baseline is a visual model which assumes that participants rely only on visual information, and an auditory model, which assumes that participants rely only on auditory information. More precisely, these baseline models assume that the participants' responses in the bimodal condition will not be different from their response in either the visual-only or the auditory-only condition. If, as we expect, the participants rely on both the auditory and the visual modalities to make decision in the bimodal condition, the optimal model would explain more variance in human responses than the visual or the auditory model do.

@@

The optimal model (as well as the auditory and visual baselines) are *normative* models. Their predictions are made about human data in the bimodal condition, but their crucial parameters (i.e., variances associated with the visual and auditory modalities) are derived from data in the unimodal conditions.
In addition to these normative models, we consider a *descriptive* model. It is formally identical to the normative optimal model (Equation 1), except that the parameters are fit to actual responses in the bimodal condition. If the referential task induces sub-optimality (due, for instance, to the arbitrary nature of the sound-object association), then the descriptive model should explain more variance than the optimal model does.

@@

There are three possible ways human responses can deviate from optimality. These scenarios are illustrated in Figure\ \@ref(fig:subOptim), and are as follows:

@@

The percentage of participants who relied exclusively on either the visual modality or the auditory modality was `r 100*(only_visual_N_3+only_sound_N_3)/(only_visual_N_3+only_sound_N_3+both_N_3)`%, which is closer to the percentage of Experiment 2, except that now almost all of them relied on the auditory modality (`r 100*(1-exlusive_prop_3)`%). For ease of comparison, Table \@ref(tab:exclusive) provides a summary of the numbers across the three experiments.

@@

When identifying a spoken word under uncertainty, one often needs to make the most of the available cues. Some previous work studied optimal behavior under uncertainty from the auditory input only [e.g., @clayard08; @feldman2009], and others studied optimality under multimodal uncertainty in auditory speech and visual facial features [e.g. @bejjanki2011]. The current work explored, for the first time, the case of word identification under uncertainty in speech (word form) and the visual *referent*. More specifically, we conducted an ideal observer analysis of the task whereby a model provided predictions about how information from each modality should be combined in an optimal fashion. The predictions of the model were tested in a series of three experiments where instances of both the form and the meaning were ambiguous with respect to their category membership only (Experiment 1), when instances of the form were perturbed with additional background noise (Experiment 2), and when instances of the referent were perturbed with additional visual noise (Experiment 3).

@@

The second case of sub-optimality is related to how participants weighted the cues from the visual and the auditory modalities in a noisy context. In contrast to Experiment 1 where the combination was indistinguishable from the optimal prediction, results of Experiment 2 and 3 which both involved background noise in one modality, showed that participants had a systematic preference for the other (non-noisy) modality. From previous empirical studies, we know that when the speech signal is degraded, people tend to compensate by relying more on other sources of information such as the accompanying visual cues (i.e., lip movements) or the semantic/syntactic context [see @mattys12 for a review]. However, and generally speaking, these studies do not differentiate between an optimal compensatory strategy (i.e., relying more on the alternative source while using all information still available in the distorted signal), and a sub-optimal strategy (i.e., relying more on the alternative source while ignoring at least some of the information still available in the distorted signal). The formal approach followed in this paper allowed us to tease apart these two possibilities, and the analysis supports the sub-optimal compensatory strategy: The preference for the non-noisy modality is above and beyond what can be explained by the relative reliability alone, meaning that the participants tend to ignore at least part of the information still available in the noisy modality.

@@

Though we only tested adults in this paper, the problem of word recognition under uncertainty, as well as the need to make the most of ambiguous cues, is a particularly pressing issue for children. In fact, whereas adults are mostly faced with uncertainty in the *input*, children have to deal with the additional uncertainty that results from their early unrefined *representations* of both phonological and semantic categories. For example, upon hearing a noisy instance of "bee", adults may have to decide whether the speaker intended to say "pea" or "bee", but children can additionally be uncertain whether "bee" is a different word from "pee" (as opposed to, say, a valid within category variation), especially if these similar sounding words are newly learned [@stager1997; @Merriman91; @Creel2012; @Swingley2016; @white2008b]. Though similar word form representation can be shown to be differentiated under some circumstances [e.g., @yoshida2009], this differentiation is still not mature enough and is probably noisier than the adult-like representation and/or encoded with lower confidence [see @Swingley2007].

@@

Thus, uncertainty in the representation associated with one modality (e.g., a bee and a fly) can be mitigated through the possibly more differentiated representations associated with the other modality (e.g., the sound "bee" is acoustically different from the sound "fly"). That being said, a multi-modal cue combination strategy might help children not only recognize an individual word instance, but also refine the underlying phonological and semantic representations in the process. Previous research in early word learning has---whether implicitly or explicitly---largely treated the process of learning form and of learning meaning as independent. However, the developmental data reviewed above shows that children do not wait to have completed the acquisition of form to start learning meanings, and that both form and meaning representations develop, rather, in a parallel fashion. A few studies pointed to the possibility of an interaction between sound and meaning in early acquisition. For instance, @waxman1995 showed that labeling various objects with the same name helps infants form the broad semantic category [but see @sloutsky2003]. Vice versa, @yeung09 showed that pairing similar sounds with different objects help infants pay attention to subtle phonological contrasts. The present study proposes a first step towards a formal framework where isolated accounts of sound-meaning interaction in development can be unified and further explored. <!--For example, one could imagine that, initially, visual and auditory categories have relatively large noise variances, and that development consists in reducing the values of the variances through a mutually constraining process as further multimodal data accumulates. --->

One limitation of this work is that we used simplified stimuli. For the auditory modality, we used speech categories that varied along a single acoustic dimension. While this dimension might be sufficient to recognize  words in our specific case, in general the speech signal may be more complex, varying along several acoustic/phonetic dimensions. Additionally, these dimensions may be highly variable due to various kinds of speaker and context differences. The same thing can be said about the referential stimuli. Here we used a continuum along a single morph dimension in order to construct a multimodal input where the auditory and visual components have symmetrical properties. Though such morph is not the exact visual variability that people would encounter in their daily lives, it allowed us to precisely test the role of auditory and visual information in the cue combination process. Parameterizing semantic dimensions is a notoriously difficult problem, but morphs have been used in previous research as a reasonable proxy [@freedman2001; @havy2016; @sloutsky2004]. 
It is an open question as to whether people use the same strategy in controlled laboratory conditions, as in more naturalistic settings where they have to deal with various levels of variability. However an answer to this question is likely to involve a multifaceted research approach, involving---besides laboratory experiments---analyses of corpora with a more realistic multimodal input [e.g., @fourtassi2014b; @harwath2016; @roy2015].

# Conclusion 
This work studied the mechanism of word identification under uncertainty in both the word form and the word referent. To our knowledge, this is the first study that performs an ideal observer analysis of this task. We found people to be near optimal in their cue combination: They weighted each modality according to its relative reliability. However, they also showed patterns of sub-optimality especially when the stimuli were perturbed with additional background noise. Though the present study did not directly address the issue of early word learning, it provides a framework where developmental questions can also be investigated. For instance, future work should explore whether children, like adults, use probabilistic cues from both the auditory and the visual input to recognize ambiguous words,  the extent to which they combine these cues in an optimal fashion, and whether these combination help them with refining their early phonological and semantic representations.
